---
title: "Variation by Climate"
format: html
---

**Purpose:** Quantify the effect of interannual variation in climatic conditions (water availability) on spatialy variability in headwater flow regimes. 

```{r, include = FALSE}
library(tidyverse)
library(mapview)
library(sf)
library(ggpubr)
library(fasstr)
library(dygraphs)
library(htmlwidgets)
library(knitr)
library(RColorBrewer)
library(scales)
library(GGally)
library(DT)
library(TSdist)
library(ggridges)
library(ggh4x)
library(viridis)
library(daymetr)
library(scico)
library(cetcolor)
library(plotly)
library(cowplot)
library(hydroEvents)
library(dtw)
library(nhdplusTools)
library(glue)
library(hydroGOF)
library(rptR)
library(moments)
library(fmsb)
library(terra)
library(ggspatial)
library(lme4)
library(MuMIn)
library(lmerTest)
library(ggfx)
library(riverdist)
```



## Data

Site information
```{r}
siteinfo <- read_csv("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Data/EcoDrought_SiteInformation.csv")
siteinfo_sp <- st_as_sf(siteinfo, coords = c("long", "lat"), crs = 4326)
```

Little g's (headwater sites)
```{r}
dat_clean <- read_csv("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Qualitative/LittleG_data_clean.csv")
```

Big G's (reference gages)
```{r}
dat_clean_big <- read_csv("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Qualitative/BigG_data_clean.csv")
```


Water availability. In western, snowmelt-dominated basins, annual (water year) total yield is strongly related to summer total yield. But in eastern, rain-dominated basins, the relationship is much weaker, suggesting "faster" response to climate forcing in rain-dominated basins. Use total summer yield (percentile) as metric of interannual variation in water availability.
```{r}
wateravail <- read_csv("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Qualitative/BigG_wateravailability_annual.csv") %>%
  filter(!is.na(totalyield), !is.na(totalyield_sum)) %>%
  group_by(site_name) %>%
  mutate(tyz_perc = percentile(totalyield_z),
         tyz_sum_perc = percentile(totalyield_sum_z)) %>%
  mutate(tyz_perc = ifelse(is.na(tyz_perc), 0, tyz_perc),
         tyz_sum_perc = ifelse(is.na(tyz_sum_perc), 0, tyz_sum_perc))

wateravail %>% 
  filter(basin != "Piney River") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River", basin)) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", "Flathead", "Yellowstone River", "Snake River", "Donner Blitzen"))) %>%
  ggplot(aes(x = tyz_perc, y = tyz_sum_perc)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  xlab("Total annual yield (percentile)") + ylab("Total summer yield (percentile)") + 
  facet_wrap(~basin) + theme_bw()
```

Order sites
```{r}
wborder <- c("West Brook NWIS", "West Brook Lower", "Mitchell Brook", "Jimmy Brook", "Obear Brook Lower", "West Brook Upper", "West Brook Reservoir", "Sanderson Brook", "Avery Brook", "West Whately Brook")
paineorder <- c("Paine Run 10", "Paine Run 08", "Paine Run 07", "Paine Run 06", "Paine Run 02", "Paine Run 01")
stauntorder <- c("Staunton River 10", "Staunton River 09", "Staunton River 07", "Staunton River 06", "Staunton River 03", "Staunton River 02")
flatorder <- c("BigCreekLower", "LangfordCreekLower", "LangfordCreekUpper", "Big Creek NWIS", "BigCreekUpper", "HallowattCreekLower", "NicolaCreek", "WernerCreek", "Hallowat Creek NWIS", "CoalCreekLower", "CycloneCreekLower", "CycloneCreekMiddle", "CycloneCreekUpper", "CoalCreekMiddle", "CoalCreekNorth", "CoalCreekHeadwaters", "McGeeCreekLower", "McGeeCreekTrib", "McGeeCreekUpper")
yellorder <- c("Shields River Valley Ranch", "Deep Creek", "Crandall Creek", "Buck Creek", "Dugout Creek", "Shields River ab Dugout", "Lodgepole Creek", "EF Duck Creek be HF", "EF Duck Creek ab HF", "Henrys Fork")
snakeorder <- c("Spread Creek Dam", "Rock Creek", "NF Spread Creek Lower", "NF Spread Creek Upper", "Grizzly Creek", "SF Spread Creek Lower", "Grouse Creek", "SF Spread Creek Upper", "Leidy Creek Mouth")
donnerorder <- c("Fish Creek NWIS", "Donner Blitzen ab Fish NWIS", "Donner Blitzen nr Burnt Car NWIS", "Donner Blitzen ab Indian NWIS")
```


Get catchment shapefiles
```{r results='hide'}
sheds_list <- list()
myfiles <- list.files(path = "C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/", pattern = ".shp")
for (i in 1:length(myfiles)) {
  sheds_list[[i]] <- st_read(paste("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/", myfiles[i], sep = ""))
}
sheds <- do.call(rbind, sheds_list) %>% 
  mutate(site_id = ifelse(site_id == "SP01", "SP07", ifelse(site_id == "SP07", "SP01", site_id))) %>%
  left_join(siteinfo)
#mapview(sheds %>% arrange(desc(area_sqmi)), alpha.regions = 0.2)
```


## Median summer flow

Plot maps of median summer flow across years/networks

::: panel-tabset
### West Brook
```{r fig.width = 9, fig.height = 5}
bas <- "West Brook"
orderr <- wborder
wtryrs <- c(2020:2024)

# sheds and network
mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Mass_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "WBR",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Mass_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# get lakes
lakes <- get_waterbodies(AOI = siteinfo_sp %>% filter(site_name == "West Brook NWIS"), buffer = 10000)
lakes <- lakes %>% filter(gnis_name %in% c("Northampton Reservoir Upper", "Northampton Reservoir"))
#st_write(lakes, "C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Study area maps/lakes/lakes_WestBrook.shp")

# little g points
mylittleg <- siteinfo_sp %>% filter(site_name %in% wborder) %>% mutate(site_name = factor(site_name, levels = wborder))


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9))

# calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# calculate long-term median of summer median flow, to center color scheme
annmed_big <- tempdat_big %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()
annmed_big <- median(annmed_big$medlogYield)

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))


# plot it
plotlist_wb <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_wb[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      #scale_fill_gradientn(colours = rev(hcl.colors(100, "Zissou1")), limits = range(annmed$medlogYield), na.value = "grey") +
      #scale_fill_gradientn(colours = cet_pal(100, name = "l16"), limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = lakes, color = "white", fill = "lightskyblue1", linewidth = 0.7) +
      geom_sf(data = lakes, color = "royalblue4", fill = "lightskyblue1", linewidth = 0.5) +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = Inf, y = Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = 1.15, hjust = 1.05, size = 5)})
}
ggarrange(plotlist = plotlist_wb, common.legend = TRUE, legend = "right")
```


### Paine Run
```{r fig.width = 6, fig.height = 5}
bas <- "Paine Run"
orderr <- paineorder
wtryrs <- c(2019:2022)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Shen_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "PA_10FL",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Shen_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# hillshade
myrast <- rast("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Data/Spatial data/Elevation/Shenandoah_DEM_10m_nc.tif")
myrast <- mask(crop(myrast, mysheds), mysheds)
slo <- terrain(myrast, "slope", unit = "radians") 
asp <- terrain(myrast, "aspect", unit = "radians")
hill <- shade(slope = slo, aspect = asp, angle = 40, direction = 270)
hilldf <- as.data.frame(hill, xy = TRUE)

# little g points
mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr)


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_pa <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_pa[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = -Inf, y = Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = 1.15, hjust = -0.05, size = 5)})
}
ggarrange(plotlist = plotlist_pa, common.legend = TRUE, legend = "right")
```

### Staunton River
```{r fig.width = 6, fig.height = 5}
bas <- "Staunton River"
orderr <- stauntorder
wtryrs <- c(2019:2022)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Shen_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "SR_10FL",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Shen_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# little g points
mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr)

# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_st <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_st[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = Inf, y = Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = 1.15, hjust = 1.05, size = 5)})
}
ggarrange(plotlist = plotlist_st, common.legend = TRUE, legend = "right")
```


### Flathead
```{r fig.height = 5, fig.width = 11}
bas <- "Flathead"
orderr <- flatorder
wtryrs <- c(2018:2024)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Flat_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "NFF",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Flat_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# get lakes
lakes <- get_waterbodies(AOI = siteinfo_sp %>% filter(site_name == "North Fork Flathead River NWIS"), buffer = 100000)
lakes <- st_transform(lakes, crs(mysheds))
lakes <- st_intersection(lakes, st_as_sf(mysheds))
lakes <- lakes %>% filter(gnis_name %in% c("Moose Lake", "Mud Lake"))

# points
mylittleg <- siteinfo_sp %>% filter(site_name %in% flatorder) %>% mutate(site_name = factor(site_name, levels = flatorder))
st_geometry(mylittleg)[mylittleg$site_name == "BigCreekUpper"] <- st_point(c(-114.31506, 48.57672))
st_geometry(mylittleg)[mylittleg$site_name == "HallowattCreekLower"] <- st_point(c(-114.31914, 48.57256))


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_fl <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_fl[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = Inf, y = Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = 1.15, hjust = 1.05, size = 5)})
}
ggarrange(plotlist = plotlist_fl, common.legend = TRUE, legend = "right")
```


### Yellowstone
```{r fig.height = 7, fig.width = 10}
bas <- "Shields River"
orderr <- yellorder
wtryrs <- c(2016:2020,2022:2024)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Shields_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id %in% c("SRS", "DU01"),]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Shields_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# points
mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% mutate(site_name = factor(site_name, levels = orderr)) #%>% filter(subbasin != "Duck Creek")


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_ye <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_ye[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = -Inf, y = -Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = -0.15, hjust = -0.05, size = 5)})
}
ggarrange(plotlist = plotlist_ye, common.legend = TRUE, legend = "right", nrow = 2, ncol = 4)
```


### Snake
```{r fig.height = 7, fig.width = 10}
bas <- "Snake River"
orderr <- snakeorder
wtryrs <- c(2016:2024)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Snake_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "SP11",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Snake_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# get lakes
lakes <- get_waterbodies(AOI = siteinfo_sp %>% filter(site_name == "Spread Creek Dam"), buffer = 100000)
lakes <- st_transform(lakes, crs(mysheds))
lakes <- st_intersection(lakes, st_as_sf(mysheds))
lakes <- lakes %>% filter(gnis_name %in% c("Leidy Lake"))

# points
mylittleg <- siteinfo_sp %>% filter(site_name %in% snakeorder) %>% mutate(site_name = factor(site_name, levels = snakeorder))
# edit geometry to reduce overlap
st_geometry(mylittleg)[mylittleg$site_name == "SF Spread Creek Lower NWIS"] <- st_point(c(-110.32226, 43.76118))
st_geometry(mylittleg)[mylittleg$site_name == "NF Spread Creek Lower"] <- st_point(c(-110.3199, 43.766533))
st_geometry(mylittleg)[mylittleg$site_name == "Grizzly Creek"] <- st_point(c(-110.23289, 43.77433))
st_geometry(mylittleg)[mylittleg$site_name == "NF Spread Creek Upper"] <- st_point(c(-110.23405, 43.77227))
st_geometry(mylittleg)[mylittleg$site_name == "SF Spread Creek Upper"] <- st_point(c(-110.31475, 43.73661))


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_sn <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_sn[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = -Inf, y = -Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = -0.15, hjust = -0.05, size = 5)})
}
ggarrange(plotlist = plotlist_sn, common.legend = TRUE, legend = "right")
```


### Donner Blitzen
```{r fig.height = 5.5, fig.width = 5.5}
bas <- "Donner Blitzen"
orderr <- c(donnerorder, "Donner Blitzen River nr Frenchglen NWIS")
wtryrs <- c(2016:2024)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Oreg_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "DBF",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Oreg_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)


# points
mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% mutate(site_name = factor(site_name, levels = donnerorder))
# edit geometry to reduce overlap
st_geometry(mylittleg)[mylittleg$site_name == "Donner Blitzen ab Fish NWIS"] <- st_point(c(-118.84315, 42.75476))
st_geometry(mylittleg)[mylittleg$site_name == "Fish Creek"] <- st_point(c(-118.83173, 42.75911))


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)


# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup() %>%
  bind_rows(tempdat_big %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup())

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_db <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_db[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      annotate("label", x = -Inf, y = -Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = -0.15, hjust = -0.05, size = 5)})
}
ggarrange(plotlist = plotlist_db, common.legend = TRUE, legend = "right")
```
:::


## Distribution and variability

Create function to plot interannual distribution of streamflow values, annual distributions of streamflow values, and scatterplot plot of spatial streamflow heterogeneity by water availability
```{r}
portfolioplot <- function(bas, orderr, type = c("interann", "annual", "scatter"), wtryrs) {
  # filter data
  tempdat <- dat_clean %>% 
    filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% wtryrs) %>%
    mutate(site_name = factor(site_name, levels = orderr))
  nsites <- length(unique(tempdat$site_name))
  tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
  varbig <- tempdat_big %>% 
    group_by(WaterYear) %>% 
    summarize(bigrange = range(logYield)[2]-range(logYield)[1],
              bigsd = sd(logYield),
              bigvar = var(logYield)) %>%
    ungroup()

  # calculate total summer water availability from reference gage
  #summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
  #wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)


  # calculate relative variation
  pedat <- tempdat %>% 
    group_by(basin, WaterYear) %>% 
    summarize(nsites = length(unique(site_name)),
              littlerange = range(logYield)[2]-range(logYield)[1],
              littlesd = sd(logYield),
              littlevar = var(logYield)) %>% 
    ungroup() %>% 
    left_join(varbig) %>%
    mutate(pe_range = ((littlerange-bigrange)/bigrange)*100,
           pe_sd = ((littlesd-bigsd)/bigsd),
           pe_var = ((littlevar-bigvar)/bigvar)) %>%
    left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_sum_perc)) %>% 
    filter(WaterYear %in% wtryrs) %>%
    mutate(wylab = substr(WaterYear, 3, 4))

  # interannual
  pint <- ggplot() + 
    geom_density(data = tempdat, aes(x = logYield, y = ..scaled.., color = site_name, fill = site_name), size = 0.8) +
    scale_color_manual(values = cet_pal(nsites, name = "i1")) +
    scale_fill_manual(values = alpha(cet_pal(nsites, name = "i1"), 0.1)) +
    geom_density(data = tempdat_big, aes(x = logYield, y = ..scaled..), color = "grey40", fill = alpha("grey40", 0.2), size = 0.8) +
    xlab("Summer log(Yield, mm/day)") + ylab("Scaled density") +
    theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                       axis.text = element_text(color = "black"), legend.position = "none")

  # annual
  pann <- ggplot() +
    geom_density(data = tempdat, aes(x = logYield, y = ..scaled.., color = site_name, fill = site_name), size = 0.8) +
    scale_color_manual(values = cet_pal(nsites, name = "i1")) +
    scale_fill_manual(values = alpha(cet_pal(nsites, name = "i1"), 0.1)) +
    geom_density(data = tempdat_big, aes(x = logYield, y = ..scaled..), color = "grey40", fill = alpha("grey40", 0.2), size = 0.8) +
    xlab("Summer log(Yield, mm/day)") + ylab("Scaled density") +
    theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                       axis.text = element_text(color = "black"), legend.position = "none") +
    facet_wrap(~factor(WaterYear, levels = as.numeric(unlist(pedat %>% arrange(totalyield_sum_z) %>% select(WaterYear)))))

  # portfolio effect by water availability
  ppew <- pedat %>% 
    ggplot(aes(x = tyz_sum_perc, y = pe_range, label = wylab)) + 
    geom_abline(slope = 0, intercept = 0, linetype = 2) +
    geom_smooth(method = "lm", color = "black") + 
    geom_point(shape = 21, fill = "skyblue1", size = 3) +
    geom_text(vjust = -0.7, color = "grey40") +
    xlab("Water availability") + ylab("Relative difference in range") +
    theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text = element_text(color = "black")) +     
    guides(color = guide_legend(title="Water year")) + xlim(0,100) + ylim(-100,450)

  if(type == "interannual") { return(pint) }
  if(type == "annual") { return(pann) }
  if(type == "scatter") { return(ppew) }
  if(type == "table") { return(pedat) }
}
```

Generate plots
```{r, include=FALSE}
# West Brook
p1 <- egg::ggarrange(portfolioplot(bas = "West Brook", orderr = wborder, type = "interannual", wtryrs = c(2020:2024)),
               portfolioplot(bas = "West Brook", orderr = wborder, type = "scatter", wtryrs = c(2020:2024)),
               nrow = 2)
p2 <- portfolioplot(bas = "West Brook", orderr = wborder, type = "annual", wtryrs = c(2020:2024))
denwb <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))

# Paine
p1 <- egg::ggarrange(portfolioplot(bas = "Paine Run", orderr = paineorder, type = "interannual", wtryrs = c(2019:2022)),
               portfolioplot(bas = "Paine Run", orderr = paineorder, type = "scatter", wtryrs = c(2019:2022)),
               nrow = 2)
p2 <- portfolioplot(bas = "Paine Run", orderr = paineorder, type = "annual", wtryrs = c(2019:2022))
denpa <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))

# Staunton
p1 <- egg::ggarrange(portfolioplot(bas = "Staunton River", orderr = stauntorder, type = "interannual", wtryrs = c(2019:2022)),
               portfolioplot(bas = "Staunton River", orderr = stauntorder, type = "scatter", wtryrs = c(2019:2022)),
               nrow = 2)
p2 <- portfolioplot(bas = "Staunton River", orderr = stauntorder, type = "annual", wtryrs = c(2019:2022))
denst <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))

# Flathead
p1 <- egg::ggarrange(portfolioplot(bas = "Flathead", orderr = flatorder, type = "interannual", wtryrs = c(2018:2024)),
               portfolioplot(bas = "Flathead", orderr = flatorder, type = "scatter", wtryrs = c(2018:2024)),
               nrow = 2)
p2 <- portfolioplot(bas = "Flathead", orderr = flatorder, type = "annual", wtryrs = c(2018:2024))
denfl <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))

# Yellowstone
p1 <- egg::ggarrange(portfolioplot(bas = "Shields River", orderr = yellorder, type = "interannual", wtryrs = c(2016:2020,2022:2024)),
               portfolioplot(bas = "Shields River", orderr = yellorder, type = "scatter", wtryrs = c(2016:2020,2022:2024)),
               nrow = 2)
p2 <- portfolioplot(bas = "Shields River", orderr = yellorder, type = "annual", wtryrs = c(2016:2020,2022:2024))
denye <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))

# Snake River
p1 <- egg::ggarrange(portfolioplot(bas = "Snake River", orderr = snakeorder, type = "interannual", wtryrs = c(2016:2024)),
               portfolioplot(bas = "Snake River", orderr = snakeorder, type = "scatter", wtryrs = c(2016:2024)),
               nrow = 2)
p2 <- portfolioplot(bas = "Snake River", orderr = snakeorder, type = "annual", wtryrs = c(2016:2024))
densn <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))

# Donner Blitzen
p1 <- egg::ggarrange(portfolioplot(bas = "Donner Blitzen", orderr = donnerorder, type = "interannual", wtryrs = c(2016:2024)),
               portfolioplot(bas = "Donner Blitzen", orderr = donnerorder, type = "scatter", wtryrs = c(2016:2024)),
               nrow = 2)
p2 <- portfolioplot(bas = "Donner Blitzen", orderr = donnerorder, type = "annual", wtryrs = c(2016:2024))
dendb <- ggarrange(p1, p2, ncol = 2, widths = c(0.7,1))
```

::: panel-tabset
##### West Brook
```{r fig.width=9, fig.height=5}
denwb
```
##### Paine Run
```{r fig.width=9, fig.height=5}
denpa
```
##### Staunton River
```{r fig.width=9, fig.height=5}
denst
```
##### Flathead
```{r fig.width=9, fig.height=5}
denfl
```
##### Yellowstone
```{r fig.width=9, fig.height=5}
denye
```
##### Snake River
```{r fig.width=9, fig.height=5}
densn
```
##### Donner Blitzen
```{r fig.width=9, fig.height=5}
dendb
```
:::


## Model heterogeneity

How does regional water availability affect spatial streamflow heterogeneity (relative difference range of streamflow values, headwaters vs. reference gages)?

Set color palette
```{r}
mycols <- brewer.pal(7, "Set2")
```


### Range diff.

FLOW REGIME: Relative difference in the range of summer flow values across all little g's compared to range of flow values at big G.

Some data manipulation
```{r}
# define basin-years to keep
keepyears <- bind_rows(
  tibble(basin = "West Brook", WaterYear = c(2020:2024)),
  tibble(basin = "Paine Run", WaterYear = c(2019:2022)),
  tibble(basin = "Staunton River", WaterYear = c(2019:2022)),
  tibble(basin = "Flathead", WaterYear = c(2018:2024)),
  tibble(basin = "Shields River", WaterYear = c(2016:2020,2022:2024)),
  tibble(basin = "Snake River", WaterYear = c(2016:2024)),
  tibble(basin = "Donner Blitzen", WaterYear = c(2016:2024)),
) %>% mutate(keep = "yes")


# calculate relative difference in range as above
mydat <- dat_clean %>% 
  filter(Month %in% c(7:9)) %>%
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1]) %>% 
  ungroup() %>% 
  left_join(dat_clean_big %>% 
              filter(Month %in% c(7:9)) %>% 
              group_by(basin, WaterYear) %>% 
              summarize(bigrange = range(logYield)[2]-range(logYield)[1]) %>%
              ungroup()) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange)*100) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 

mydat
```

Show simple linear model summary, weighted by number of sites
```{r}
summary(lm(pe_range ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r}
par(mar = c(9,4,0.5,0.5))
boxplot(pe_range ~ basin, data = mydat, ylab = "Spatial streamflow heterogeneity", xlab = "", las = 2, col = mycols)
```


#### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(pe_range ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = pe_range, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = pe_range, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial streamflow heterogeneity (percent difference)") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot1 <- randiffplot


# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

Use *rptR* to decompose variance in spatial streamflow heterogeneity into fixed (regional water availability) and random (basin) effect sources:
```{r message=FALSE}
rmod <- rptGaussian(pe_range ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat, grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
rmod$mod
plot(rmod, grname = "basin")
plot(rmod, grname = "Fixed")
```

The strong effect of basin is largely driven by Staunton and Donner-Blitzen Rivers. Without those basins, random effect variance shrinks and fixed effect variance expands considerably. Much of this pattern is likely due to low sample sizes (few headwater gages in these basins, particularly in the Donner-Blitzen)
```{r message=FALSE}
rmod <- rptGaussian(pe_range ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat %>% filter(!basin %in% c("Staunton River", "Donner-Blitzen River")), grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
```


#### Sensitivity analysis

How sensitive are these results to the choice of data (log) transformation, per comments from John Hammond during USGS internal review?

Summary: While the among-network mean effect of water availability on heterogeneity was less significant using un-logged discharge data (p > 0.05), this appeared to be driven by the highly non-normal distribution of raw discharge values, as the reviewer indicated. Importantly, the magnitude of the effect is similar and the qualitative results are unchanged (negative slope). 

Data manipulation: use un-logged/raw Yield data to calculation spatial heterogeneity, also use CV rather than ranges
```{r}
mydat2 <- dat_clean %>% 
  filter(Month %in% c(7:9)) %>%
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(Yield_mm)[2]-range(Yield_mm)[1],
            littlecv = sd(Yield_mm)/mean(Yield_mm)) %>% 
  ungroup() %>% 
  left_join(dat_clean_big %>% 
              filter(Month %in% c(7:9)) %>% 
              group_by(basin, WaterYear) %>% 
              summarize(bigrange = range(Yield_mm)[2]-range(Yield_mm)[1],
                        bigcv = sd(Yield_mm)/mean(Yield_mm)) %>%
              ungroup()) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange)*100,
         pe_cv = ((littlecv-bigcv)/bigcv)*100) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 
```


##### Range, un-logged

Fit the mixed effects model
```{r}
mydat2 <- mydat2 %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(pe_range ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat2)
summary(lmod)
```

Plot
```{r}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat2$z_tyz_sum_perc), max(mydat2$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat2$z_tyz_sum_perc, "scaled:scale")) + attr(mydat2$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat2 %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat2 %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat2, aes(x = tyz_sum_perc, y = pe_range, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial streamflow heterogeneity (percent difference)") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed")
```


##### CV of summer flows

Relative difference in the CV of summer flow values across all little g's compared to CV of flow values at big G

Fig the mmixed effects model
```{r}
mydat2 <- mydat2 %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(pe_cv ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat2)
summary(lmod)
```

Plot
```{r}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat2$z_tyz_sum_perc), max(mydat2$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat2$z_tyz_sum_perc, "scaled:scale")) + attr(mydat2$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat2 %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat2 %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat2, aes(x = tyz_sum_perc, y = pe_cv, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial streamflow heterogeneity (percent difference)") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed")

```


##### CV of time series distance

FLOW REGIME: Spatial variation (CV) in the time series distance between each little g and corresponding big G (CV of TSdist)

Some data manipulation
```{r}
# define basin-years to keep
keepyears <- bind_rows(
  tibble(basin = "West Brook", WaterYear = c(2020:2024)),
  tibble(basin = "Paine Run", WaterYear = c(2019:2022)),
  tibble(basin = "Staunton River", WaterYear = c(2019:2022)),
  tibble(basin = "Flathead", WaterYear = c(2018:2024)),
  tibble(basin = "Shields River", WaterYear = c(2016:2020,2022:2024)),
  tibble(basin = "Snake River", WaterYear = c(2016:2024)),
  tibble(basin = "Donner Blitzen", WaterYear = c(2016:2024)),
) %>% mutate(keep = "yes")


# calculate mean and cv of among little g time series distances
mydat <- dat_clean %>% 
  filter(Month %in% c(7:9)) %>%
  left_join(dat_clean_big %>% 
              filter(Month %in% c(7:9)) %>% 
              select(basin, WaterYear, Yield_mm, logYield) %>% 
              rename("bigYield_mm" = "Yield_mm", "biglogYield" = "logYield")) %>%
  group_by(basin, site_name, WaterYear) %>%
  summarize(tsdist = EuclideanDistance(Yield_mm, bigYield_mm),
            tsdist_logvals = EuclideanDistance(logYield, biglogYield)) %>%
  ungroup() %>%
  group_by(basin, WaterYear) %>%
  summarize(mean_tsd = mean(tsdist),
            mean_tsdlog = mean(tsdist_logvals),
            cv_tsd = sd(tsdist) / mean(tsdist),
            cv_tsdlog = sd(tsdist_logvals) / mean(tsdist_logvals),
            nsites = n()) %>%
  ungroup() %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 

mydat
# 
# mydat %>% 
#   ggplot(aes(x = tyz_sum_perc, y = mean_tsd)) + 
#   geom_point(aes(fill = basin, size = nsites), shape = 21) + 
#   scale_fill_brewer(palette = "Set2") +
#   theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
#   ylab("Mean time series distance") + xlab("Water availability (percentile)") +
#   geom_smooth(method = "lm") +
#   guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
#   geom_abline(intercept = 0, slope = 0, linetype = "dashed")
#   
# mydat %>% 
#   ggplot(aes(x = tyz_sum_perc, y = mean_tsdlog)) + 
#   geom_point(aes(fill = basin, size = nsites), shape = 21) + 
#   scale_fill_brewer(palette = "Set2") +
#   theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
#   ylab("Mean time series distance (log vals)") + xlab("Water availability (percentile)") +
#   geom_smooth(method = "lm") +
#   guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
#   geom_abline(intercept = 0, slope = 0, linetype = "dashed")
#   
# mydat %>% 
#   ggplot(aes(x = tyz_sum_perc, y = cv_tsd)) + 
#   geom_point(aes(fill = basin, size = nsites), shape = 21) + 
#   scale_fill_brewer(palette = "Set2") +
#   theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
#   ylab("CV time series distance") + xlab("Water availability (percentile)") +
#   geom_smooth(method = "lm") +
#   guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
#   geom_abline(intercept = 0, slope = 0, linetype = "dashed")
#   
# mydat %>% 
#   ggplot(aes(x = tyz_sum_perc, y = cv_tsdlog)) + 
#   geom_point(aes(fill = basin, size = nsites), shape = 21) + 
#   scale_fill_brewer(palette = "Set2") +
#   theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
#   ylab("CV time series distance (log vals)") + xlab("Water availability (percentile)") +
#   geom_smooth(method = "lm") +
#   guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
#   geom_abline(intercept = 0, slope = 0, linetype = "dashed")
```

Show simple linear model summary, weighted by number of sites
```{r}
summary(lm(cv_tsdlog ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r}
par(mar = c(9,5,0.5,0.5))
boxplot(cv_tsdlog ~ basin, data = mydat, ylab = "Spatial CV time series distance\n(log values)", xlab = "", las = 2, col = mycols)
```


Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cv_tsdlog ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cv_tsdlog, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cv_tsdlog, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial CV time series distance") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot2 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

Use *rptR* to decompose variance in spatial streamflow heterogeneity into fixed (regional water availability) and random (basin) effect sources:
```{r message=FALSE}
rmod <- rptGaussian(cv_tsdlog ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat, grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
rmod$mod
plot(rmod, grname = "basin")
plot(rmod, grname = "Fixed")
```

The strong effect of basin is largely driven by Staunton and Donner-Blitzen Rivers. Without those basins, random effect variance shrinks and fixed effect variance expands considerably. Much of this pattern is likely due to low sample sizes (few headwater gages in these basins, particularly in the Donner-Blitzen)
```{r message=FALSE}
rmod <- rptGaussian(cv_tsdlog ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat %>% filter(!basin %in% c("Staunton River", "Donner-Blitzen River")), grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
```


## Paper figs

### West Brook

2020/2021 summer flow distributions
```{r fig.width=2.75, fig.height=5}
bas <- "West Brook"
orderr <- wborder
wtryrs <- c(2020,2021)

# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% wtryrs) %>%
  mutate(site_name = factor(site_name, levels = orderr))

nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

# annual
pann <- ggplot() +
  geom_density(data = tempdat, aes(x = logYield, y = ..scaled.., color = site_name, fill = site_name), size = 0.8) +
  scale_color_manual(values = cet_pal(nsites, name = "i1")) +
  scale_fill_manual(values = alpha(cet_pal(nsites, name = "i1"), 0.1)) +
  geom_density(data = tempdat_big, aes(x = logYield, y = ..scaled..), color = "grey40", fill = alpha("grey40", 0.2), size = 0.8) +
  xlab("Summer log(Yield, mm/day)") + ylab("Density") +
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
                     axis.text = element_text(color = "black"), legend.position = "none",
                     strip.background = element_blank(), strip.text = element_blank()) +
  facet_wrap(~factor(WaterYear, levels = as.numeric(unlist(pedat %>% arrange(totalyield_sum_z) %>% select(WaterYear)))), ncol = 1)

print(pann)


# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_summerdensity_WB_annual.jpg", width = 2.75, height = 5, units = "in", res = 1000)
# pann
# dev.off()
```

2020/2021 summer flow distributions
```{r fig.width=4, fig.height=6}
bas <- "West Brook"
orderr <- wborder
wtryrs <- c(2020,2021)

# sheds and network
mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Mass_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "WBR",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Mass_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# get lakes
lakes <- get_waterbodies(AOI = siteinfo_sp %>% filter(site_name == "West Brook NWIS"), buffer = 10000)
lakes <- lakes %>% filter(gnis_name %in% c("Northampton Reservoir Upper", "Northampton Reservoir"))

# little g points
mylittleg <- siteinfo_sp %>% filter(site_name %in% wborder) %>% mutate(site_name = factor(site_name, levels = wborder))


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_wb <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_wb[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = lakes, color = "white", fill = "lightskyblue1", linewidth = 0.7) +
      geom_sf(data = lakes, color = "royalblue4", fill = "lightskyblue1", linewidth = 0.5) +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      theme_void() + scale_x_continuous(expand = c(0.003, 0.003)) + scale_y_continuous(expand = c(0.002, 0.002))
      #annotate("label", x = Inf, y = Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = 1.15, hjust = 1.05, size = 5)
      })
}
p <- ggarrange(plotlist = plotlist_wb, common.legend = TRUE, legend = "right", nrow = 2)
print(p)


# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_summermedianmap_WB_annual.jpg", width = 4, height = 6, units = "in", res = 1000)
# p
# dev.off()
```


### Snake River

2021/2022 summer flow distributions
```{r fig.width=2.75, fig.height=5}
bas <- "Snake River"
orderr <- snakeorder
wtryrs <- c(2021,2022)

# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% wtryrs) %>%
  mutate(site_name = factor(site_name, levels = orderr))

nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

# annual
pann <- ggplot() +
  geom_density(data = tempdat, aes(x = logYield, y = ..scaled.., color = site_name, fill = site_name), size = 0.8) +
  scale_color_manual(values = cet_pal(nsites, name = "i1")) +
  scale_fill_manual(values = alpha(cet_pal(nsites, name = "i1"), 0.1)) +
  geom_density(data = tempdat_big, aes(x = logYield, y = ..scaled..), color = "grey40", fill = alpha("grey40", 0.2), size = 0.8) +
  xlab("Summer log(Yield, mm/day)") + ylab("Density") +
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
                     axis.text = element_text(color = "black"), legend.position = "none",
                     strip.background = element_blank(), strip.text = element_blank()) +
  facet_wrap(~factor(WaterYear, levels = as.numeric(unlist(pedat %>% arrange(totalyield_sum_z) %>% select(WaterYear)))), ncol = 1)

print(pann)


# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_summerdensity_Snake_annual.jpg", width = 2.75, height = 5, units = "in", res = 1000)
# pann
# dev.off()
```

2022/2022 summer flow distributions
```{r fig.width=4, fig.height=6}
bas <- "Snake River"
orderr <- snakeorder
wtryrs <- c(2021,2022)

mysheds <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Watersheds/Snake_Watersheds.shp")
mysheds <- mysheds[mysheds$site_id == "SP11",]
mynet <- vect("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/EcoDrought-Analysis/Watershed Delineation/Streams/Snake_Streams.shp")
crs(mynet) <- crs(mysheds)
mynet <- crop(mynet, mysheds)

# get lakes
lakes <- get_waterbodies(AOI = siteinfo_sp %>% filter(site_name == "Spread Creek Dam"), buffer = 100000)
lakes <- st_transform(lakes, crs(mysheds))
lakes <- st_intersection(lakes, st_as_sf(mysheds))
lakes <- lakes %>% filter(gnis_name %in% c("Leidy Lake"))

# points
mylittleg <- siteinfo_sp %>% filter(site_name %in% snakeorder) %>% mutate(site_name = factor(site_name, levels = snakeorder))
# edit geometry to reduce overlap
st_geometry(mylittleg)[mylittleg$site_name == "SF Spread Creek Lower NWIS"] <- st_point(c(-110.32226, 43.76118))
st_geometry(mylittleg)[mylittleg$site_name == "NF Spread Creek Lower"] <- st_point(c(-110.3199, 43.766533))
st_geometry(mylittleg)[mylittleg$site_name == "Grizzly Creek"] <- st_point(c(-110.23289, 43.77433))
st_geometry(mylittleg)[mylittleg$site_name == "NF Spread Creek Upper"] <- st_point(c(-110.23405, 43.77227))
st_geometry(mylittleg)[mylittleg$site_name == "SF Spread Creek Upper"] <- st_point(c(-110.31475, 43.73661))


# filter data
tempdat <- dat_clean %>% 
  filter(basin == bas, !is.na(logYield), Month %in% c(7:9)) %>%
  mutate(site_name = factor(site_name, levels = orderr))
nsites <- length(unique(tempdat$site_name))
tempdat_big <- dat_clean_big %>% filter(basin == bas, !is.na(logYield), Month %in% c(7:9), WaterYear %in% unique(tempdat$WaterYear))

  # calculate variability for reference gage
varbig <- tempdat_big %>% 
  group_by(WaterYear) %>% 
  summarize(bigrange = range(logYield)[2]-range(logYield)[1],
            bigsd = sd(logYield),
            bigvar = var(logYield)) %>%
  ungroup()

# calculate total summer water availability from reference gage
#summerflow <- tempdat_big %>% group_by(WaterYear) %>% summarize(summeryield = sum(Yield_mm), summeryield_log = log(sum(Yield_mm))) %>% ungroup()
#wateravail2 <- wateravail %>% select(basin, WaterYear, totalyield, totalyield_z)

# calculate relative variation
pedat <- tempdat %>% 
  group_by(basin, WaterYear) %>% 
  summarize(nsites = length(unique(site_name)),
            littlerange = range(logYield)[2]-range(logYield)[1],
            littlesd = sd(logYield),
            littlevar = var(logYield)) %>% 
  ungroup() %>% 
  left_join(varbig) %>%
  mutate(pe_range = ((littlerange-bigrange)/bigrange),
         pe_sd = ((littlesd-bigsd)/bigsd),
         pe_var = ((littlevar-bigvar)/bigvar)) %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield, totalyield_z, totalyield_sum, totalyield_sum_z, tyz_perc, tyz_sum_perc)) %>% 
  filter(WaterYear %in% wtryrs) %>%
  mutate(wylab = substr(WaterYear, 3, 4)) %>%
  arrange(totalyield_sum_z)
  
# calculate annual median summer flow
annmed <- tempdat %>%
  #filter(WaterYear %in% c(lowyr, highyr)) %>%
  group_by(site_name, WaterYear) %>%
  summarize(medlogYield = median(logYield, na.rm = TRUE)) %>%
  ungroup()

# arrange years by total annual yield
wtryrs_arranged <- as.numeric(unlist(pedat %>% select(WaterYear)))

# plot it
plotlist_wb <- list()
for(i in 1:length(wtryrs_arranged)) {
  mylittleg <- siteinfo_sp %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i]))
  mylittleg_sheds <- sheds %>% filter(site_name %in% orderr) %>% left_join(annmed %>% filter(WaterYear == wtryrs_arranged[i])) %>% arrange(desc(area_sqmi))
  plotlist_wb[[i]] <- local({
    i <- i
    ggplot() +
      geom_sf(data = st_as_sf(mysheds), color = "black", fill = "white", linewidth = 0.4) + 
      geom_sf(data = mylittleg_sheds, aes(fill = medlogYield), color = "black") +
      scale_fill_viridis(option = "G", direction = 1, limits = range(annmed$medlogYield), na.value = "grey") +
      geom_sf(data = st_as_sf(mynet), color = "white", linewidth = 1, lineend = "round") +
      geom_sf(data = st_as_sf(mynet), color = "royalblue4", linewidth = 0.6, lineend = "round") +
      geom_sf(data = lakes, color = "white", fill = "lightskyblue1", linewidth = 0.7) +
      geom_sf(data = lakes, color = "royalblue4", fill = "lightskyblue1", linewidth = 0.5) +
      geom_sf(data = mylittleg, shape = 21, fill = "white", size = 2) +
      labs(fill = "Median\nsummer\nlog(Yield)") + #annotation_scale() +
      coord_sf(xlim = range(st_coordinates(mylittleg_sheds)[,1]), ylim = range(st_coordinates(mylittleg_sheds)[,2])) +
      theme_bw() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none", axis.text = element_blank()) +
      theme_void() + scale_x_continuous(expand = c(0.003, 0.003)) + scale_y_continuous(expand = c(0.002, 0.002))
      #annotate("label", x = Inf, y = Inf, label = paste(pedat$WaterYear[i], " (", pedat$tyz_sum_perc[i], "%)", sep = ""), vjust = 1.15, hjust = 1.05, size = 5)
      })
}
p <- ggarrange(plotlist = plotlist_wb, common.legend = TRUE, legend = "right", nrow = 2)
print(p)


# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_summermedianmap_Snake_annual.jpg", width = 4, height = 6, units = "in", res = 1000)
# p
# dev.off()
```



::: {.content-hidden}

## Heterogeneity in flow metrics

A couple of the reviewers wanted to see the heterogeneity analysis repeated for flow metrics. See below for the above analysis applied to metrics describing summer flow magnitude, rate of change, duration, and timing. None of the relationships with annual water available are significant at alpha = 0.05. Importantly, summarizing flow regime data into metrics obscures much of the important patterns of the data

### Magnitude

MAGNITUDE: effect of water availability on *spatial variation in median summer specific discharge* (spatial CV of median summer specific discharge by basin and year)

Some data manipulation
```{r eval=FALSE}
# define basin-years to keep
keepyears <- bind_rows(
  tibble(basin = "West Brook", WaterYear = c(2020:2024)),
  tibble(basin = "Paine Run", WaterYear = c(2019:2022)),
  tibble(basin = "Staunton River", WaterYear = c(2019:2022)),
  tibble(basin = "Flathead", WaterYear = c(2018:2024)),
  tibble(basin = "Shields River", WaterYear = c(2016:2020,2022:2024)),
  tibble(basin = "Snake River", WaterYear = c(2016:2024)),
  tibble(basin = "Donner Blitzen", WaterYear = c(2016:2024)),
) %>% mutate(keep = "yes")


# calculate relative difference in range as above
mydat <- dat_clean %>%
  filter(Month %in% c(7:9)) %>%
  group_by(basin, site_name, WaterYear) %>%
  summarize(medyield = median(Yield_mm, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dat_clean_big %>% 
              filter(Month %in% c(7:9)) %>% 
              group_by(basin, WaterYear) %>% 
              summarize(medyield_big = median(Yield_mm, na.rm = TRUE)) %>%
              ungroup()) %>%
  mutate(medyield_scaled = medyield / medyield_big) %>%
  group_by(basin, WaterYear) %>%
  summarize(cvyield = sd(medyield)/mean(medyield), nsites = n()) %>%
  ungroup() %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 

mydat
```

Show simple linear model summary, weighted by number of sites
```{r eval=FALSE}
summary(lm(cvyield ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r eval=FALSE}
par(mar = c(9,4,0.5,0.5))
boxplot(cvyield ~ basin, data = mydat, ylab = "CV median summer streamflow", xlab = "", las = 2, col = mycols)
```


#### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r eval=FALSE}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cvyield ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r eval=FALSE}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4, eval=FALSE}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cvyield, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cvyield, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial CV median summer streamflow") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot3 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

Use *rptR* to decompose variance in spatial streamflow heterogeneity into fixed (regional water availability) and random (basin) effect sources:
```{r message=FALSE, eval=FALSE}
rmod <- rptGaussian(cvyield ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat, grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
rmod$mod
plot(rmod, grname = "basin")
plot(rmod, grname = "Fixed")
```

The strong effect of basin is largely driven by Staunton and Donner-Blitzen Rivers. Without those basins, random effect variance shrinks and fixed effect variance expands considerably. Much of this pattern is likely due to low sample sizes (few headwater gages in these basins, particularly in the Donner-Blitzen)
```{r message=FALSE, eval=FALSE}
rmod <- rptGaussian(cvyield ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat %>% filter(!basin %in% c("Staunton River", "Donner-Blitzen River")), grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
```


### Rate of change

RATE OF CHANGE: effect of water availability on spatial variation in *Richard-Baker flashiness index* (spatial CV RBI by basin and year)

Some data manipulation
```{r eval=FALSE}
# define basin-years to keep
keepyears <- bind_rows(
  tibble(basin = "West Brook", WaterYear = c(2020:2024)),
  tibble(basin = "Paine Run", WaterYear = c(2019:2022)),
  tibble(basin = "Staunton River", WaterYear = c(2019:2022)),
  tibble(basin = "Flathead", WaterYear = c(2018:2024)),
  tibble(basin = "Shields River", WaterYear = c(2016:2020,2022:2024)),
  tibble(basin = "Snake River", WaterYear = c(2016:2024)),
  tibble(basin = "Donner Blitzen", WaterYear = c(2016:2024)),
) %>% mutate(keep = "yes")


mydat <- dat_clean %>%
  filter(Month %in% c(7:9)) %>%
  group_by(basin, site_name, WaterYear) %>%
  summarize(ar1c = acf(Yield_mm, lag.max = 1, plot = FALSE)$acf[2],
            ar1c_log = acf(logYield, lag.max = 1, plot = FALSE)$acf[2],
            rbi = sum(abs(Yield_mm - lag(Yield_mm)), na.rm = TRUE) / sum(Yield_mm, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(basin, WaterYear) %>%
  summarize(cvar1c = sd(ar1c)/mean(ar1c), 
            cvar1c_log = sd(ar1c_log)/mean(ar1c_log), 
            cvrbi = sd(rbi)/mean(rbi),
            nsites = n()) %>%
  ungroup() %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 

mydat

ggpairs(dat_clean %>%
  filter(Month %in% c(7:9)) %>%
  group_by(basin, site_name, WaterYear) %>%
  summarize(ar1c = acf(Yield_mm, lag.max = 1, plot = FALSE)$acf[2],
            ar1c_log = acf(logYield, lag.max = 1, plot = FALSE)$acf[2],
            rbi = sum(abs(Yield_mm - lag(Yield_mm)), na.rm = TRUE) / sum(Yield_mm, na.rm = TRUE)) %>%
  ungroup() %>% select(ar1c, ar1c_log, rbi))
```

Show simple linear model summary, weighted by number of sites
```{r eval=FALSE}
summary(lm(cvrbi ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r eval=FALSE}
par(mar = c(9,4,0.5,0.5))
boxplot(cvrbi ~ basin, data = mydat, ylab = "CV Richard-Baker flashiness index", xlab = "", las = 2, col = mycols)
```


#### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r eval=FALSE}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cvrbi ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r eval=FALSE}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4, eval=FALSE}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cvrbi, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cvrbi, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial CV Richard-Baker flashiness index") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot4 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

Use *rptR* to decompose variance in spatial streamflow heterogeneity into fixed (regional water availability) and random (basin) effect sources:
```{r message=FALSE, eval=FALSE}
rmod <- rptGaussian(cvyield ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat, grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
rmod$mod
plot(rmod, grname = "basin")
plot(rmod, grname = "Fixed")
```

The strong effect of basin is largely driven by Staunton and Donner-Blitzen Rivers. Without those basins, random effect variance shrinks and fixed effect variance expands considerably. Much of this pattern is likely due to low sample sizes (few headwater gages in these basins, particularly in the Donner-Blitzen)
```{r message=FALSE, eval=FALSE}
rmod <- rptGaussian(cvyield ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat %>% filter(!basin %in% c("Staunton River", "Donner-Blitzen River")), grname = c("basin", "Fixed"), ratio = TRUE, adjusted = FALSE, nboot = 1000)
print(rmod)
```



### Duration

Duration: effect of water availability on spatial variation in *autoregressive lag-one (AR1) correlation* (spatial CV site and year specific AR1 by basin and year)

Some data manipulation
```{r eval=FALSE}
# define basin-years to keep
keepyears <- bind_rows(
  tibble(basin = "West Brook", WaterYear = c(2020:2024)),
  tibble(basin = "Paine Run", WaterYear = c(2019:2022)),
  tibble(basin = "Staunton River", WaterYear = c(2019:2022)),
  tibble(basin = "Flathead", WaterYear = c(2018:2024)),
  tibble(basin = "Shields River", WaterYear = c(2016:2020,2022:2024)),
  tibble(basin = "Snake River", WaterYear = c(2016:2024)),
  tibble(basin = "Donner Blitzen", WaterYear = c(2016:2024)),
) %>% mutate(keep = "yes")

# 
# p1 <- plot_flow_duration(dat_clean %>% filter(Month %in% c(7:9), site_name == "Avery Brook"),
#                    dates = date, values = Yield_mm, custom_months = c(7:9))
# calc_longterm_daily_stats(dat_clean %>% filter(Month %in% c(7:9), site_name == "Avery Brook"),
#                           dates = date, values = Yield_mm)
# 
# calc_flow_percentile(dat_clean %>% filter(Month %in% c(7:9), site_name == "Avery Brook"),
#                           dates = date, values = Yield_mm)

calc_longterm_percentile(dat_clean %>% filter(Month %in% c(7:9), site_name == "Avery Brook"),
                          dates = date, values = Yield_mm, months = c(7:9), percentiles = c(10,50,90))

mydat <- dat_clean %>%
  filter(Month %in% c(7:9)) %>%
  group_by(basin, site_name, WaterYear) %>%
  group_modify( ~ calc_longterm_percentile(.x, dates = date, values = Yield_mm, months = c(7:9), percentiles = c(10,50,90))) %>%
  ungroup() %>%
  group_by(basin, WaterYear) %>%
  summarize(cvP10 = sd(P10)/mean(P10), 
            cvP50 = sd(P50)/mean(P50), 
            cvP90 = sd(P90)/mean(P90),
            nsites = n()) %>%
  ungroup() %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 

mydat

```

#### 10th percentile

Show simple linear model summary, weighted by number of sites
```{r eval=FALSE}
summary(lm(cvP10 ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r eval=FALSE}
par(mar = c(9,4,0.5,0.5))
boxplot(cvP10 ~ basin, data = mydat, ylab = "CV 10th percentile", xlab = "", las = 2, col = mycols)
```


##### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r eval=FALSE}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cvP10 ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r eval=FALSE}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4, eval=FALSE}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cvP10, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cvP10, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab(expression(paste("Spatial CV ", 10^th, " percentile streamflow"))) + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot5 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

#### 50th percentile

Show simple linear model summary, weighted by number of sites
```{r eval=FALSE}
summary(lm(cvP50 ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r eval=FALSE}
par(mar = c(9,4,0.5,0.5))
boxplot(cvP50 ~ basin, data = mydat, ylab = "Spatial CV 50th percentile", xlab = "", las = 2, col = mycols)
```


##### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r eval=FALSE}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cvP50 ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r eval=FALSE}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4, eval=FALSE}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cvP50, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cvP50, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab(expression(paste("Spatial CV ", 50^th, " percentile streamflow"))) + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot6 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

#### 90th percentile

Show simple linear model summary, weighted by number of sites
```{r eval=FALSE}
summary(lm(cvP90 ~ tyz_sum_perc, data = mydat, weights = nsites))
```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r eval=FALSE}
par(mar = c(9,4,0.5,0.5))
boxplot(cvP90 ~ basin, data = mydat, ylab = "Spatial CV 90th percentile", xlab = "", las = 2, col = mycols)
```


##### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r eval=FALSE}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cvP90 ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r eval=FALSE}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4, eval=FALSE}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cvP90, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cvP90, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab(expression(paste("Spatial CV ", 90^th, " percentile streamflow"))) + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot
randiffplot7 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```

### Timing

TIMING:
Some data manipulation
```{r eval=FALSE}
# define basin-years to keep
keepyears <- bind_rows(
  tibble(basin = "West Brook", WaterYear = c(2020:2024)),
  tibble(basin = "Paine Run", WaterYear = c(2019:2022)),
  tibble(basin = "Staunton River", WaterYear = c(2019:2022)),
  tibble(basin = "Flathead", WaterYear = c(2018:2024)),
  tibble(basin = "Shields River", WaterYear = c(2016:2020,2022:2024)),
  tibble(basin = "Snake River", WaterYear = c(2016:2024)),
  tibble(basin = "Donner Blitzen", WaterYear = c(2016:2024)),
) %>% mutate(keep = "yes")



mybas <- dat_clean %>% group_by(basin, site_name) %>% summarize(n = n()) %>% ungroup() %>% select(-n)
annstats <- calc_annual_stats(data = dat_clean %>% filter(Month %in% c(7:9)),
                  dates = date, values = Yield_mm, groups = site_name, months = c(7:9), ignore_missing = TRUE)
annlows <- calc_annual_lowflows(data = dat_clean %>% filter(Month %in% c(7:9)),
                  dates = date, values = Yield_mm, groups = site_name, months = c(7:9), ignore_missing = TRUE)
mydat <- annstats %>% left_join(annlows) %>% left_join(mybas) %>%
  mutate(WaterYear = Year) %>%
  select(basin, site_name, WaterYear, Mean, Median, P10, Min_7_Day, Min_7_Day_DoY) %>%
  group_by(basin, WaterYear) %>%
  summarize(cv_Mean = sd(Mean, na.rm = TRUE) / mean(Mean, na.rm = TRUE), 
            cv_Median = sd(Median, na.rm = TRUE) / mean(Median, na.rm = TRUE), 
            cv_P10 = sd(P10, na.rm = TRUE) / mean(P10, na.rm = TRUE), 
            cv_Min_7_Day = sd(Min_7_Day, na.rm = TRUE) / mean(Min_7_Day, na.rm = TRUE), 
            cv_Min_7_Day_DoY = sd(Min_7_Day_DoY, na.rm = TRUE) / mean(Min_7_Day_DoY, na.rm = TRUE),
            nsites = n()) %>%
  ungroup() %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River", 
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run", 
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River"))) 
 
 
# # calculate relative difference in range as above
# thresh <- dat_clean %>%
#   filter(Month %in% c(7:9)) %>%
#   group_by(basin, site_name) %>%
#   group_modify( ~ calc_longterm_percentile(.x, dates = date, values = Yield_mm, months = c(7:9), percentiles = c(5,10,50))) %>%
#   ungroup()
# 
mydat <- dat_clean %>%
  filter(Month %in% c(7:9)) %>%
  left_join(thresh) %>%
  group_by(basin, site_name, WaterYear) %>%
  summarize(skewness = moments::skewness(logYield, na.rm = TRUE),
            kurtosis = moments::kurtosis(logYield, na.rm = TRUE)) %>%
  group_by(basin, WaterYear) %>%
  summarize(cv_skew = sd(skewness) / mean(skewness),
            cv_kurt = sd(kurtosis) / mean(kurtosis),
            nsites = n()) %>%
  ungroup() %>%
  left_join(wateravail %>% select(basin, WaterYear, totalyield:tyz_sum_perc)) %>%
  left_join(keepyears) %>% filter(keep == "yes") %>%
  mutate(basin = ifelse(basin == "Shields River", "Yellowstone River",
                        ifelse(basin == "Flathead", "Flathead River",
                               ifelse(basin == "Donner Blitzen", "Donner-Blitzen River", basin)))) %>%
  mutate(basin = factor(basin, levels = c("West Brook", "Staunton River", "Paine Run",
                                          "Flathead River", "Yellowstone River", "Snake River", "Donner-Blitzen River")))

mydat
```

Show simple linear model summary, weighted by number of sites
```{r eval=FALSE}
summary(lm(cv_skew ~ tyz_sum_perc, data = mydat, weights = nsites))
summary(lm(cv_kurt ~ tyz_sum_perc, data = mydat, weights = nsites))

```

Show distribution of spatial streamflow heterogeneity by basin. Lots of structuring by basin, driven in large part by Staunton and Donner-Blitzen Rivers.
```{r eval=FALSE}
par(mar = c(9,4,0.5,0.5))
boxplot(cv_skew ~ basin, data = mydat, ylab = "CV days below site-level 50th percentile", xlab = "", las = 2, col = mycols)
```


#### Fit model 

Use a random slope/random intercept mixed effects model to estimate the among-basin mean effect of regional water availability on spatial streamflow heterogeneity.

**Notes:**

* This regression is not weighted by number of headwater sites used to calculate spatial heterogeneity because variance decomposition using *rptR* does not support use of weights.
* This uses z-scored x-values...using the raw percentiles leads to model convergence issues

```{r eval=FALSE}
mydat <- mydat %>% mutate(z_tyz_sum_perc = scale(tyz_sum_perc))
lmod <- lmer(cv_kurt ~ z_tyz_sum_perc + (1 + z_tyz_sum_perc | basin), data = mydat)
summary(lmod)
```

Get p-values for global slope and intercept: (doesn't to run in Quarto...works fine in R)
```{r eval=FALSE}
coef(summary(as_lmerModLmerTest(lmod)))
```


Plot relationship between spatial streamflow heterogeneity and regional water availability with global (among-population mean) regression line. 
```{r fig.width = 5, fig.height = 4, eval=FALSE}
# get predicted values and 95% credible interval
effects_tyz <- effects::effect(term = "z_tyz_sum_perc", mod = as_lmerModLmerTest(lmod), xlevels = list(z_tyz_sum_perc = seq(min(mydat$z_tyz_sum_perc), max(mydat$z_tyz_sum_perc), length.out = 1000)), se = list(level = 0.95))
x_tyz <- as_tibble(effects_tyz) %>% mutate(tyz_sum_perc = (z_tyz_sum_perc * attr(mydat$z_tyz_sum_perc, "scaled:scale")) + attr(mydat$z_tyz_sum_perc, "scaled:center"))

# trim data for panel labesl
mydat_labs <- bind_rows(mydat %>% filter(basin == "West Brook", WaterYear %in% c(2020,2021)),
                        mydat %>% filter(basin == "Snake River", WaterYear %in% c(2021,2022))) %>%
  mutate(labs = c("b", "c", "d", "e"))

# plot 
randiffplot <- ggplot() + 
  geom_ribbon(data = x_tyz, aes(x = tyz_sum_perc, ymin = lower, ymax = upper), fill = "grey80") +
  geom_line(data = x_tyz, aes(x = tyz_sum_perc, y = fit), color = "black", size = 1.25, lineend = "round") +
  geom_point(data = mydat, aes(x = tyz_sum_perc, y = cv_skew, fill = basin, size = nsites), shape = 21) + 
  scale_fill_brewer(palette = "Set2") +
  geom_text(data = mydat_labs, aes(x = tyz_sum_perc, y = cv_skew, label = labs), size = 3) +
  theme_bw() + theme(panel.grid = element_blank(), axis.text = element_text(color = "black")) +
  ylab("Spatial CV median summer streamflow") + xlab("Water availability (percentile)") +
  guides(fill = guide_legend(title = "Basin", override.aes = list(size = 5), order = 1), size = guide_legend(title = "Number of sites", order = 2)) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1, fontface = "italic",
           label = paste("p = ", round(coef(summary(as_lmerModLmerTest(lmod)))[2,5], digits = 3)))
randiffplot + ylim(0,2.5)
# randiffplot3 <- randiffplot

# write to file
# jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_RangeDiff_allsites2.jpg", width = 4.75, height = 5, units = "in", res = 1000)
# randiffplot
# dev.off()
```


### Combined figs

```{r fig.width=8, fig.height=8, eval=FALSE}
bp1 <- egg::ggarrange(randiffplot1 + theme(legend.position = "none", axis.title.x = element_blank()), 
                      randiffplot2 + theme(axis.title.x = element_blank()), ncol = 2)

bp2 <- egg::ggarrange(randiffplot3 + theme(legend.position = "none", axis.title.x = element_blank()), 
                      randiffplot4 + theme(legend.position = "none", axis.title.x = element_blank()), 
                      randiffplot5 + theme(legend.position = "none", axis.title.x = element_blank()), 
                      randiffplot7 + theme(legend.position = "none", axis.title.x = element_blank()), 
                      nrow = 1, bottom = "Annual water availability (percentile)")

ggarrange(bp1, bp2, nrow = 2)

jpeg("C:/Users/jbaldock/OneDrive - DOI/Documents/USGS/EcoDrought/EcoDrought Working/Presentation figs/EcoD_FlowMetricsByWaterAvail.jpg", width = 8, height = 8, units = "in", res = 1000)
ggarrange(bp1, bp2, nrow = 2)
dev.off()
```


